---
title: "The World is Changing"
subtitle: "Accessible data for climate change"
author: "by Team ACE"
output: github_document
---
```{r load-packages, message = FALSE, echo=FALSE}
library(tidyverse)
library(broom)
library(leaflet) ## For leaflet interactive maps
library(sf) ## For spatial data
library(RColorBrewer) ## For color palettes
library(htmltools) ## For html
library(leafsync) ## For placing plots side by side
library(kableExtra) ## Table output
library(tidymodels) # modelling
library(readr)
```


## Introduction

Climate change is an ever increasing concern for humanity as we continue to rely on natural gas, oil, and other finite resources to power our lives. Despite efforts to reduce personal carbon footprints and switch to renewable sources of energy such as solar, wind, or hydroelectric, greenhouse gas emissions are still rising worldwide. If we as a species do not take action to switch from fossil fuels to sustainable sources of energy, we risk losing more than we can handle, from the biodiversity of many ecosystems to entire housing markets. The only way to reduce the effects of climate change is to stop using fossil fuels soon, however even though people are aware of climate change, facts alone will not engage people to take action. As such, we have taken these facts and turned them into more powerful figures to help people visualize the damage done by climate change. 

## Methodology

The data analyzed to make visualizations for this project came from a variety of sources. The majority of which, including sea level rise, city population density, and temperature came from Kaggle. All of the data analyzed in this project were numerical data, with exceptions for city names and disaster types, which were qualitative data. Datasets were combined and cleaned to create data frames that were easy to work with and create plots. For example, the US_Temp_Pop dataset combines US city population data and temperature data over time to more easily visualize how temperatures in cities with high populations  are rising. Using these datasets in combination with one another helped us answer our initial question, “How does increasing global temperatures measured by city predict the number of natural disasters in the same area?” This was accomplished by making histogram plots of natural disasters, plots of specific disasters such as wildfires per city, and creating a leaflet mapping of wildfires throughout the United States. 

Although natural disasters are striking and cause lots of acute damage when they occur, rising sea levels also pose a more subtle threat. As sea levels continue to rise in the long term, more and more habitats and coastal communities become increasingly concerned with consistent flooding and destruction of their property and ecosystems. For this reason, we also included plots of rising sea levels globally. In addition to the plot, we also created a linear regression model to interpret the rate at which global sea level rise is occurring at. The model also has the added benefit of being able to predict what the mean sea level may look like in the near future based on past data. Finally, we also analyzed global CO2 levels as an explanatory variable for the increase in global temperature rise.  

## Findings

The linear regression formulated from mean sea level rise reveals that global mean sea levels are rising at approximately .318mm/year.. At this current rate, coastal communities like Miami beaches could be completely submerged by 2100( Castro, Alejandro).  While this rate was formulated based on current data, it is likely that this rate will increase in the coming years as the rate of CO2 emissions is increasing exponentially. This is evident both in our visualization of the Mauna Loa CO2 levels, and also in our figure depicting the Global CO2 levels over the past 800 thousand years, in which we find a sharp spike in the CO2 levels in ppm at the dawn of the industrial revolution. 

In concordance with rising CO2 levels, there is also a linear increase in the average temperature of cities in the US with a population over one million. Higher temperatures increase risk to areas with drier climates as periods of heat waves last longer. This in turn increases the risk of drought in the area along with the probability of wildfires. Our line plot showing the large wildfires per year from 1990-2015 portrays an increase concurrent with rising temperatures. Most strikingly, the leaflet plot visualizing large wildfires across the US outlines that wildfires are a much more serious concern on the west coast and in Alaska compared to the east coast. 

Between rising sea levels, increasing rates of wildfires, and higher CO2 levels in our atmosphere, it is clear that as we continue to rely on fossil fuels to power our lives, we pay a higher price each day by risking more frequent and catastrophic natural disasters. 





```{r load-data, echo = FALSE, message=FALSE}
global_co2 <- readr::read_csv(file = "data/co2_data.csv")
disaster_data <- readr::read_csv(file = "data/us_disaster_declarations.csv")
```


```{r, echo = FALSE, fig.alt="Density plot of natural disasters. Shows increase from 1950 to present", warning=FALSE, message=FALSE}
disaster_data %>%
ggplot(mapping = aes(x = fy_declared)) +
  geom_density() + 
  theme(axis.text.x = element_text(angle = 45)) + 
  labs(title = "Histogram of major disasters",
       subtitle = "By disaster type",
       x = "Year",
       y = "Count") +
  scale_color_viridis_c()
```



```{r global_co2, echo = FALSE, fig.alt="Global CO2 levels (ppm) of last 800 thousand years from ice cores. Big spike at onset of industrial revolution and still rising today.", warning=FALSE, message=FALSE}
global_co2 %>%
  mutate(ageKa = Age_yrBP/1000) %>%
ggplot(mapping = aes(x = ageKa, y = CO2_ppm)) + 
  geom_line() + 
  scale_x_reverse() +
  theme_bw() + 
   labs(title = "Global CO2 levels",
      subtitle = "Over the past 800ka",
      x = "Age (Ka BP)",
      y = "CO2 levels (ppm)")
```

```{r recent_co2, echo = FALSE, fig.alt="Mauna Loa CO2 levels (ppm) of last 60 years. Shows seasonal variability and distinct exponential increase.", warning=FALSE, message=FALSE}
global_co2 %>%
  mutate(Date = as.Date(Date)) %>%
ggplot(mapping = aes(x = Date, y = co2_ppm)) + 
  geom_line() +
  geom_smooth(color = "red", se = FALSE) +
  theme_bw() + 
  scale_x_date(date_labels = "%Y %b %d") + 
  labs(title = "Mauna Loa CO2 levels",
       subtitle = "Since 1958",
       x = "Date",
       y = "CO2 levels (ppm)") 
```


## Presentation

Our presentation can be found [here](presentation/presentation.html).

## Data

<!-- Include a citation for your data here. -->
<!-- See http://libraryguides.vu.edu.au/c.php?g=386501&p=4347840 for guidance on proper citation for datasets. -->
<!-- If you got your data off the web, make sure to note the retrieval date. -->

US Temperature Data

File name: US_temps.csv

[Source](https://www.kaggle.com/berkeleyearth/climate-change-earth-surface-temperature-data)

US Disaster Data
- File name: us_disaster_declarations.csv
- [Source](https://www.kaggle.com/headsortails/us-natural-disaster-declarations/version/72)

Wildfire Data
- File name: FW_Veg_Rem_Combined.csv
- [Source](https://www.kaggle.com/capcloudcoder/us-wildfire-data-plus-other-attributes)

Sea Level Data
- File name: sealevel.csv
- [Source](https://www.kaggle.com/kkhandekar/global-sea-level-1993-2021)

US city population density Data
- File name: uscitypopdensity.csv
- [Source](https://www.kaggle.com/mmcgurr/us-city-population-densities)

Global CO2 data
- File name: co2_data.csv
- [Source](https://keelingcurve.ucsd.edu/permissions-and-data-sources/)

## References

Castro, Alejandro. “Miami Underwater.” ArcGIS StoryMaps, Esri, 4 Dec. 2020, https://storymaps.arcgis.com/stories/80d2465cd19c405f933e9afcfe1fdf83. 

- [Climate Image](https://climate.nasa.gov/solutions/adaptation-mitigation/)
- [Water Scarcity](https://www.worldwildlife.org/threats/water-scarcity)


