---
title: "Project proposal"
author: "Team ACE"
output:
  github_document: default
  html_document: default
---


```{r load-packages, message = FALSE}
library(tidyverse)
library(broom)
library(leaflet) ## For leaflet interactive maps
library(sf) ## For spatial data
library(RColorBrewer) ## For colour palettes
library(htmltools) ## For html
library(leafsync) ## For placing plots side by side
library(kableExtra) ## Table output
library(tidymodels) # modelling
library(readr)
```

## 1. Introduction
Our research question is "How does increasing global temperatures measured by city predict the number of natural disasters in the same area?" Our data for global temperatures by city was found using kaggle's climate change database. Our data ranges from 1750 to 2015, and has 32 columns representing variables with information regarding min, max, and average land temperature. 


Our sub topics include:

1. Global temp and wildfires
2. Global temp and droughts
3. Global temp and flooding
4. Global temp and hurricanes




## Load Data
```{r load-data-US disaster declarations, message=FALSE}
disaster_data <- readr::read_csv(file = "../data/us_disaster_declarations.csv")
glimpse(disaster_data)
```

```{r load-data-wildfires, message=FALSE}
wildfires <- readr::read_csv(file = "../data/FW_Veg_Rem_Combined.csv")
glimpse(wildfires)
```

```{r load-data-US_temps, message=FALSE}
US_temp <- readr::read_csv(file = "../data/US_temps.csv")
glimpse(US_temp)
```

```{r load-data-sea_lvl, message=FALSE}
sea_lvl <- readr::read_csv(file = "../data/sealevel.csv")
glimpse(sea_lvl)
```

```{r load-data-US_pop, message=FALSE}
US_pop <- readr::read_csv(file = "../data/uscitypopdensity.csv")
glimpse(US_pop)
US_pop$Population_2016 <- US_pop$"2016 Population" # create a new column/variable
US_pop$"2016 Population" <- NULL # remove the column
```




## 2. Data

1. US Temperature - [Source](https://www.kaggle.com/berkeleyearth/climate-change-earth-surface-temperature-data)
2. US Disasters - [Source](https://www.kaggle.com/headsortails/us-natural-disaster-declarations/version/72)
3. US Wildfires - [Source](https://www.kaggle.com/capcloudcoder/us-wildfire-data-plus-other-attributes)
4. Sea Level - [Source](https://www.kaggle.com/kkhandekar/global-sea-level-1993-2021)
5. US City Population - [Source](https://www.kaggle.com/mmcgurr/us-city-population-densities)






## 3. Data analysis plan
The first analysis we will do is to visualize how global temperatures can be used as a predictor for the number of wildfires in a given year within a specific area.

The first analysis we will do is to visualize how global temperatures can be used as a predictor for the number of wildfires in a given year within a specific area. 


```{r merge-us_temp+pop, echo=FALSE}
US_temp_pop <- left_join(x = US_temp,
                         y = US_pop,
                         by = "City")
```


### STATISTICAL METHODS

Some statistical methods we should use in our project are linear regressions to determine how correlated the two datasets are and some statistical significance tests to see if our conclusions are reliable. 


### RESULTS NEEDED

For our question to be answered, we would need a significantly correlated dataset that would either show a positive association or negative association between global temperature and wildfire probability. 

## Examples of relevant graphs
```{r histogram-of-wildfires}
ggplot(data = wildfires,
       mapping = aes(x = disc_pre_year)) + 
  #geom_histogram(binwidth = 1) + 
  geom_density() +
  labs(title = "Histogram of wildfires",
       subtitle = "From 1991 to 2015",
       x = "Year",
       y = "Count")
```

```{r san_fran_temps}
US_temp %>%
  filter(City == "San Francisco") %>%
    ggplot(mapping = aes(x = dt, 
                         y = AverageTemperature)) + 
    geom_smooth() + 
    labs(title = "Temperature in San Fransisco, Ca",
         subtitle = "Over the past sesquintennial",
         x = "Year",
         y = "Average Temperature")

```

```{r disaster-hist}
disaster_data %>%
  filter(incident_type %in% c("Coastal Storm", "Drought", "Earthquake", "Fire", "Flood","Freezing", "Hurricane", "Mud/Landslide", "Severe Ice Storm","Severe Storm(s)", "Snow", "Tornado", "Tsunami", "Typhoon",
"Volcano")) %>%
  ggplot(mapping = aes(x = fy_declared)) +
    geom_histogram(binwidth = 10) + 
    facet_wrap( ~ incident_type) +
    theme(axis.text.x = element_text(angle = 45)) + 
    labs(title = "Histogram of major disasters",
         subtitle = "By disaster type",
         x = "Year",
         y = "Count")
```

```{r large-wildfires}
wildfires %>%
  filter(fire_size_class == c("D", "E", "F", "G")) %>%
  ggplot(mapping = aes(x = disc_pre_year, 
                       y = fire_size)) + 
    geom_smooth() + 
    labs(title = "Large wildfires by year",
         subtitle = "Wildfires of class D, E, F, and G, size of 100 acres or more",
         x = "Year of wildfire",
         y = "Wildfire size (acres)")
  
```

```{r}
ggplot(data = sea_lvl,
       mapping = aes(x = Year, 
                     y = GMSL_noGIA)) + 
  geom_point() + 
  geom_smooth() +
  labs(title = "Global mean sea level rise",
       subtitle = "Relative to 2006",
       x = "Year",
       y = "Global Mean Sea Level (mm)")
```

```{r temp-large-cities}
US_temp_pop %>% 
  filter(Population_2016 >= 1000000) %>% # cities with over 1 million population
  select(dt, AverageTemperature, City, State, Population_2016) %>%
  ggplot(mapping = aes(x = dt, 
                       y = AverageTemperature)) + 
    geom_smooth() + 
    labs(title = "Temperature in Large Cities in the US",
         subtitle = "Over the past sesquintennial, filtered for cities with pop > 1000000",
         x = "Year",
         y = "Average Temperature") + 
    facet_wrap( ~ City)

```


Need to do: remove letters from lat and long in `US_temp_pop`.
```{r}
US_temp_pop <- US_temp_pop %>%
  mutate(Latitude = str_remove_all(Latitude, "[NESW]"),
         Longitude = str_remove_all(Longitude, "[NESW]"))
US_temp_pop$Longitude <- as.numeric(US_temp_pop$Longitude)
US_temp_pop$Latitude <- as.numeric(US_temp_pop$Latitude)
US_temp_pop <- US_temp_pop %>%
  mutate(Longitude = Longitude * -1) 
```


2 leaflet maps:
 - wildfires colored by size, have grouped circles
 - map showing cities and populations, colored by average increase in temperature (would need linear model I think)
```{r leaflet-map, eval = FALSE}
leaflet(data = US_temp_pop) %>%
  addTiles() %>%
  setView(lng = -97, 
          lat = 39, 
          zoom = 4) %>%
  addCircleMarkers(lng = Longitude, 
                   lat = Latitude, 
                   label = City, 
                   clusterOptions = markerClusterOptions())
```



