---
title: "Project proposal"
author: "Team ACE"
output:
  github_document: default
  html_document: default
---


```{r load-packages, message = FALSE}
library(tidyverse)
library(broom)
```

## 1. Introduction
Our research question is "How does increasing global temperatures measured by city predict the number of natural disasters in the same area?" Our data for global temperatures by city was found using kaggle's climate change database. Our data ranges from 1750 to 2015, and has 32 columns representing variables with information regarding min, max, and average land temperature. 


Our sub topics include:

1. Global temp and wildfires
2. Global temp and droughts
3. Global temp and flooding
4. Global temp and hurricanes




## Load Data
```{r load-data-US disaster declarations, message=FALSE}
disaster_data <- readr::read_csv(file = "../data/us_disaster_declarations.csv")
glimpse(disaster_data)
```

```{r load-data-wildfires, message=FALSE}
wildfires <- readr::read_csv(file = "../data/FW_Veg_Rem_Combined.csv")
glimpse(wildfires)
```

```{r load-data-US_temps, message=FALSE}
US_temp <- readr::read_csv(file = "../data/US_temps.csv")
glimpse(US_temp)
```

```{r load-data-sea_lvl, message=FALSE}
sea_lvl <- readr::read_csv(file = "../data/sealevel.csv")
glimpse(sea_lvl)
```

```{r load-data-US_pop, message=FALSE}
US_pop <- readr::read_csv(file = "../data/uscitypopdensity.csv")
glimpse(US_pop)
```



## 2. Data
Some of our data sources include:
1. Csv file from kaggle with global temp by city 
2. Csv file from kaggle with wildfire info
3. CSV file from kaggle on US disaster declarations (includes info on floods and hurricanes)
and some more... add later





## 3. Data analysis plan
The first analysis we will do is to visualize how global temperatures can be used as a predictor for the number of wildfires in a given year within a specific area. Continuing off this theme, we will also analyze how global sea levels rise with rising global temperature as well. To give emphasis on why this is a problem, we will also include population data to see how much impact sea level rise and natural disasters have on a city. 


```{r merge-us_temp+pop}
#US_temp_pop <- left_join(x = US_temp,
#                         y = US_pop,
#                         by = "City") #%>%
#                #filter(population > 200000)

```


### STATISTICAL METHODS

Some statistical methods we should use in our project are linear regressions to determine how correlated the two datasets are and some statistical significance tests to see if our conclusions are reliable. 


### RESULTS NEEDED

For our question to be answered, we would need a significantly correlated dataset that would either show a positive association or negative association between global temperature and wildfire probability. 

## Example of relevant graphs
```{r histogram-of-wildfires}
ggplot(data = wildfires,
       mapping = aes(x = disc_pre_year)) + 
  geom_histogram(binwidth = 1) + 
  labs(title = "Histogram of wildfires",
       subtitle = "From 1991 to 2015",
       x = "Year",
       y = "Count")
```

```{r san_fran_temps}
US_temp %>%
  filter(City == "San Francisco") %>%
    ggplot(mapping = aes(x = dt, 
                         y = AverageTemperature)) + 
    geom_smooth() + 
    labs(title = "Temperature in San Fransisco, Ca",
         subtitle = "Over the past sesquintennial",
         x = "Year",
         y = "Average Temperature")

```

```{r disaster-hist}
disaster_data %>%
  filter(incident_type %in% c("Coastal Storm", "Drought", "Earthquake", "Fire", "Flood","Freezing", "Hurricane", "Mud/Landslide", "Severe Ice Storm","Severe Storm(s)", "Snow", "Tornado", "Tsunami", "Typhoon",
"Volcano")) %>%
  ggplot(mapping = aes(x = fy_declared)) +
    geom_histogram(binwidth = 30) + 
    facet_wrap( ~ incident_type) +
    theme(axis.text.x = element_text(angle = 45))
```





