---
theme: jekyll-theme-minimal
title: "The World is Changing"
subtitle: "An Analysis of Rising Temperatures and All The Fun That Comes with It"
author: "Team ACE <br> Austin Gilbride, Chenemi Maji, Evan Saltman"
institute: "Bates College"
date: "April 11, 2022"
output:
  xaringan::moon_reader:
    css: xaringan-themer-ace.css
    lib_dir: libs
    nature:
      ratio: "16:9"
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false
---

```{r load-packages, include = FALSE}
# Add any additional packages you need to this chunk
library(tidyverse)
library(tidymodels)
library(palmerpenguins)
library(knitr)
library(xaringanthemer)
library(broom)
library(leaflet) ## For leaflet interactive maps
library(sf) ## For spatial data
library(RColorBrewer) ## For color palettes
library(htmltools) ## For html
library(leafsync) ## For placing plots side by side
library(kableExtra) ## Table output
library(tidymodels) # modelling
library(readr)
```

```{r setup, include=FALSE}
# For better figure resolution
knitr::opts_chunk$set(fig.retina = 3, dpi = 300, fig.width = 6, fig.asp = 0.618, out.width = "80%")
```

```{r xaringan-themer, include=FALSE, warning=FALSE}
style_mono_dark(
  base_color = "#cbf7ed",
  title_slide_text_color = "#000000",
  header_font_google = google_font("Josefin Sans"),
  text_font_google   = google_font("Montserrat", "300", "300i"),
  code_font_google   = google_font("Fira Mono")
  )
```


```{r load-data, include=FALSE}
disaster_data <- readr::read_csv(file = "../data/us_disaster_declarations.csv")
wildfires <- readr::read_csv(file = "../data/FW_Veg_Rem_Combined.csv")
US_temp <- readr::read_csv(file = "../data/US_temps.csv")
sea_lvl <- readr::read_csv(file = "../data/sealevel.csv")
US_pop <- readr::read_csv(file = "../data/uscitypopdensity.csv")
global_co2 <- readr::read_csv(file = "../data/co2_data.csv")
```

```{r, echo=FALSE}
style_xaringan(

  title_slide_background_image = "img/globe_image_map.jpeg",
  title_slide_text_color = "#FF0000",
  title_slide_background_color = "#595959",
  title_slide_background_size = "contain"
  )
```

---
class: center, middle

# How do rising global temperatures broadly affect the frequency and severity of natural disasters? 


```{r include=FALSE, echo=FALSE}
US_pop$Population_2016 <- US_pop$"2016 Population" # create a new column/variable
US_pop$"2016 Population" <- NULL # remove the column
US_temp_pop <- left_join(x = US_temp,
                         y = US_pop,
                         by = "City")
```

class: inverse, center, middle

# Global temperatures and their relationship to natural disasters

---
```{r temp-large-cities, echo=FALSE}
US_temp_pop %>% 
  filter(Population_2016 >= 1000000) %>% # cities with over 1 million population
  select(dt, AverageTemperature, City, State, Population_2016) %>%
  ggplot(mapping = aes(x = dt, 
                       y = AverageTemperature)) + 
    geom_smooth() + 
    labs(title = "Temperature in Large Cities in the US",
         subtitle = "Over the past sesquintennial, filtered for cities with pop > 1000000",
         x = "Year",
         y = "Average Temperature") + 
    facet_wrap( ~ City)

```




```{r large-wildfires, echo=FALSE}
lg_wildfires <- wildfires %>%
  filter(fire_size_class %in% c("D", "E", "F", "G"))
lg_wildfires %>%
  ggplot(mapping = aes(x = disc_pre_year, 
                       y = fire_size)) + 
    geom_smooth() + 
    labs(title = "Large wildfires by year",
         subtitle = "Wildfires of class D, E, F, and G, size of 100 acres or more",
         x = "Year of wildfire",
         y = "Wildfire size (acres)")
```


---
#How has the global sea level increased? 
.pull-left[
* Sea level has continued to increase at faster and faster rates.

* In concordance with increasing CO2 levels
]
.pull-right[
```{r sea level rise cleaning, echo=FALSE}
sea_lvl <- sea_lvl %>%
  group_by(Year) %>%
  mutate(avg_GMSL_per_year = mean(GMSL_GIA))
```

```{r sea level rise mapping, echo=FALSE}
ggplot(data = sea_lvl,
       mapping = aes(x = Year, 
                     y = GMSL_noGIA)) + 
  geom_point() + 
  geom_line(aes(y = avg_GMSL_per_year, color = "#edf8b1")) +
  geom_point(aes(y = avg_GMSL_per_year, color = "yellow")) +
  geom_smooth() +
  labs(title = "Global mean sea level rise",
       subtitle = "Relative to 2006",
       x = "Year",
       y = "Global Mean Sea Level (mm)") +
  scale_color_viridis_d()

```
]

---
## Linear Model of Global Mean Sea level rise and US temperature and Population

.pull-left[
* A linear model helps us predict mean global sea level

* GMSL_GIA = (0.318)year + 2004.781
]
.pull-right[
```{r linear-regression-modeling-on-SLR, echo=FALSE}
US_temp_pop <- US_temp_pop %>%
  filter(Population_2016 >= 1000000)%>%
  m1 <- linear_reg() %>% # Select Model Type
         set_engine("lm") %>% # Set engine
         fit(AverageTemperature ~ City, data = US_temp_pop) #code kept crashing so switched to sea_lvl instead


m1 <- linear_reg() %>% # Select Model Type
      set_engine("lm") %>% # Set engine
      fit(Year ~ GMSL_GIA , data = sea_lvl)
summary(m1)
```

```{r sea-level-regression-mapping, echo=FALSE}
sea_lvl %>%
ggplot(aes(x = Year, y = GMSL_GIA)) +
  geom_point(alpha = 0.4) +
  geom_smooth(method = "lm", se = FALSE) +
  geom_line(aes(y = avg_GMSL_per_year)) +
  labs(title = "Yearly Increase in global mean sea level rise",
      subtitle = "A linear model",
      x = "Year",
      y = "Global Mean Sea Level")

```
]

---
# Visualizing the Increase in Disasters Throughout the US. 

```{r disaster-hist, echo=FALSE}
disaster_data %>%
  filter(incident_type %in% c("Coastal Storm", "Drought", "Earthquake", "Fire", "Flood","Freezing", "Hurricane", "Mud/Landslide", "Severe Ice Storm","Severe Storm(s)", "Snow", "Tornado", "Tsunami", "Typhoon",
"Volcano")) %>%
  ggplot(mapping = aes(x = fy_declared)) +
    geom_histogram(binwidth = 10) + 
    #geom_density() + 
    facet_wrap( ~ incident_type) +
    theme(axis.text.x = element_text(angle = 45)) + 
    labs(title = "Histogram of major disasters",
         subtitle = "By disaster type",
         x = "Year",
         y = "Count") +
  scale_color_viridis_c()
```


#Leaflet Mapping of Large WildFires in the US
---
.pull-left[
 - wildfires have caused $X amount of damage to communities
 - can be prevented with $X amount of preventative work
]
.pull-rght[
```{r, echo=FALSE}
US_temp_pop <- US_temp_pop %>%
  mutate(Latitude = str_remove_all(Latitude, "[NESW]"),
         Longitude = str_remove_all(Longitude, "[NESW]"))
US_temp_pop$Longitude <- as.numeric(US_temp_pop$Longitude)
US_temp_pop$Latitude <- as.numeric(US_temp_pop$Latitude)
US_temp_pop <- US_temp_pop %>%
  mutate(Longitude = Longitude * -1) 
```

```{r wildfires-by-size, echo=FALSE}
wf_pal <- colorFactor(palette = "YlOrRd", domain = lg_wildfires$fire_size_class)
lg_wildfires %>%
  leaflet() %>%
  addTiles() %>%
  setView(lng = -97, 
          lat = 39, 
          zoom = 4) %>%
  addCircles(lng = ~lg_wildfires$longitude, 
             lat = ~lg_wildfires$latitude, 
             label = ~lg_wildfires$stat_cause_descr, ~lg_wildfires$disc_pre_year, 
             radius = ~lg_wildfires$wf_rad_mi,
             color = wf_pal(lg_wildfires$fire_size_class),
             stroke = FALSE,
             fillOpacity = 0.7,
             #clusterOptions = markerClusterOptions()
             )
```
]

---
#Relating global CO2 levels with increasing temperature

```{r recent_co2, echo=FALSE}
global_co2 %>%
  mutate(Date = as.Date(Date)) %>%
ggplot(mapping = aes(x = Date, y = co2_ppm)) + 
  geom_line() +
  geom_smooth(color = "red", se = FALSE) +
  theme_bw() + 
  scale_x_date(date_labels = "%Y %b %d") + 
  labs(title = "Mauna Loa CO2 levels",
       subtitle = "Since 1958",
       x = "Date",
       y = "CO2 levels (ppm)") 
```

---
#CO2 level increase over the past 800 thousand years 
.pull-right[
- Earth has historically been in this range
<- Since the industrial revolution we are now here
]

.pull-left[
```{r global_co2, echo=FALSE}
global_co2 %>%
  mutate(ageKa = Age_yrBP/1000) %>%
ggplot(mapping = aes(x = ageKa, y = CO2_ppm)) + 
  geom_line() + 
  scale_x_reverse() +
  theme_bw() + 
   labs(title = "Global CO2 levels",
      subtitle = "Over the past 800ka",
      x = "Age (Ka BP)",
      y = "CO2 levels (ppm)")
```
]

---
#Thank you and Questions 

